Developing a Visual Studio Code Language Support Extension for the Snail Programming Language

### 2023-06-14 paper.tex
feature: draft section 4.1

\usepackage[title]{appendix}
\usepackage{bera} % not sure I love this font for this document
\usepackage{cite}
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{listings}
\usepackage{wrapfig}
\usepackage{xcolor}

@@

\begin{wrapfigure}{c}{0.3\textwidth}
    \begin{center}
        \includegraphics[height=0.3\textheight]{directory-structure.png}
        \caption{The typical directory structure of a VSCode language support extension}
    \end{center}
    \label{fig:directory-structure}
\end{wrapfigure}

@@

\begin{wrapfigure}{r}{0.3\textwidth}
    \begin{center}
        \includegraphics[width=0.3\textwidth]{lsp-basic.png}
        \caption{A development tool will communicate with a language server to display advanced diagnostics. The language server protocol standardizes this communication.}
    \end{center}
    \label{fig:lsp-basic}
\end{wrapfigure}

A language server is a tool used to provide language-specific editing features such as advanced autocomplete, go-to definition, or automatic error checking \cite{Microsoft_2022a}. Many IDEs today utilize language servers. A development tool will launch a separate language server process to run in the background while a developer writes code. The development tool will then send a \emph{request} to the language server for advanced diagnostics, and the language server will \emph{respond} with the information requested by the development tool. The development tool will then display the diagnostics to the developer. This process is displayed in figure Figure \ref{fig:lsp-basic}.

@@

\begin{wrapfigure}{l}{0.3\textwidth}
    \begin{center}
        \includegraphics[width=0.3\textwidth]{no-lsp-theory.png}
        \caption{Test}
    \end{center}
    \label{fig:no-lsp-theory}
\end{wrapfigure}

@@

>>>>

\usepackage[title]{appendix}
\usepackage{bera} % not sure I love this font for this document
\usepackage{cite}
\usepackage{caption}
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{listings}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{xcolor}

@@

\begin{wrapfigure}{c}{0.3\textwidth}
    \begin{center}
        \includegraphics[height=0.3\textheight]{directory-structure.png}
        \caption{The typical directory structure of a VSCode language support extension}
        \label{fig:directory-structure}
    \end{center}
\end{wrapfigure}

@@

\begin{wrapfigure}{r}{0.3\textwidth}
    \begin{center}
        \includegraphics[width=0.3\textwidth]{language-server.png}
        \caption{A development tool will communicate with a language server to display advanced diagnostics. The language server protocol standardizes this communication.}
        \label{fig:language-server}
    \end{center}
\end{wrapfigure}

A language server is a tool used to provide language-specific editing features such as advanced autocomplete, go-to definition, or automatic error checking \cite{Microsoft_2022a}. Many IDEs today utilize language servers. A development tool will launch a separate language server process to run in the background while a developer writes code. The development tool will then send a \emph{request} to the language server for advanced diagnostics, and the language server will \emph{respond} with the information requested by the development tool. The development tool will then display the diagnostics to the developer. This process is displayed in figure Figure \ref{fig:language-server}.

@@

@@

The LSP devines a number of common interactions between a development tool and a langauge server. For example, a development tool might send the \lstinline{DidchangeTextDocument} notification when a development tool detects that a text document has been edited. A language server might validate this text document and send any notable diagnostics back to the development tool.

With an established standard for communication, both development tools and language servers can streamline their communications systems. Additionally, a developer producing one of these tools or features can rely on a consistent standard to make the development process easier. This benefit is also illustrated in Figure \ref{fig:lsp-theory}.

\begin{figure}
    \begin{subfigure}[c][][c]{0.45\textwidth}
        \begin{center}
            \includegraphics[width=\textwidth]{no-lsp-theory.png}
            \caption{Without the Language Server Protocol, development tool and language server communication can get messy.}
            \label{fig:no-lsp-theory}
        \end{center}
    \end{subfigure}
    \hspace{0.05\textwidth}
    \begin{subfigure}[c][][c]{0.45\textwidth}
        \begin{center}
            \includegraphics[width=\textwidth]{lsp-theory.png}
            \caption{The Language Server Protocol streamlines the communication between development tools and language servers.}
            \label{fig:lsp-theory}
        \end{center}
    \end{subfigure}
    \caption{Demonatrating the benefit of VS Code's Language Server Protocol.}
    \label{fig:lsp}
\end{figure}

In VS Code extensions, a language server (using LSP) can provide a number of different features for a programming language. By implementing support for the \lstinline{textDocument/hover} request, an exstension could display additional documentation about code when a user hovers their cursor over some text. By implementing support for the \lstinline{widnow/showMessage} request, an extension can display an error message in VS Code's UI when code fails to build. With the \lstinline{textDocument/publishDiagnostics} notification, an extension can validate that a program will run before a user actually tries to run it. For a full list of LSP specificatons, see the documentation \cite{Microsoft_2023i}.
